### 2014-05-09 latex/paper1.tex
starting paper1

\documentclass[12pt, oneside, letterpaper, fleqn]{article}

\usepackage{bobianite}

\pagestyle{fancy}
\headheight 1.6cm
\fancyhead[L]{Bryance Oyang\\HPS 160}

\begin{document}
\begin{center}
\textbf{title}
\end{center}

\nocite{*}
\bibliographystyle{unsrt}
\bibliography{bib_paper1}

\end{document}