### 2014-05-09 latex/outline_paper1.tex
wrote zeeman stuff

\documentclass[12pt, oneside, letterpaper, fleqn]{article}

\usepackage{bobianite}

\begin{document}
\begin{center}
\textbf{Outline}
\end{center}
\begin{itemize}

\item physics and technology for experiment had progressed enough that
it was possible for the first time to investigate the structure of
matter. physics evolved from being concerned with the behavior of things
to what makes up those things, what are the constituents of matter,
energy, light, sound.

\item Zeeman
\begin{itemize}
\item a good amount of Zeeman's paper was devoted to investigating the
structure of atoms
\item he put heavy emphasis on what his experimental results implied
about the structure of matter
\item zeeman discusses his correspondence with Lorentz: a model for the
atom and how the model was in agreement with experiment
\item with the model he estimated the charge to mass ratio of the
charged particles in the atom
\item faraday had previously tried this, but did not have precise
experimental apparatus that zeeman had
\item zeeman performed several versions of the experiment: in both he
observed a widening of the spectral line of atoms when a magnetic field
was applied
\item lorentz and zeeman's model of the atom predicted the line would
split into 3 lines. the 3 lines blurred together would look as if the
line widened
\end{itemize}

\end{itemize}
\end{document}